High-Performance Data Analysis on JANUS using Apache Spark

### 2013-12-18 paper.tex
adding pagerank overview

>>>>

\subsubsection{PageRank}   
Since Spark system is an in-memory cluster computing system, the power of such a system should be exploited by applications that runs several iterations on the same data.

As an example of iterative algorithms the \emph{Page Rank} algorithm was chosen.
The original pagerank algorithm was developed by Google websearch engine which enabled them to rank the webpages in their search engine.

\emph{Data set}:
The source of the dataset is DBPedia, a crowd sourced community that extracts structured information from wikipedia and makes it available on the web. \citep{[??]} There are different formats of wikipedia datasets available online. For our experiments we chose the wikipedia pagelink dataset \citep{[??]} , this contains
internal links between the DBPedia instances extracted from the internal links of the wikipedia articles making it apt for the any datamining related analysis including Pagerank algorithm. Therefore this dataset was chosen for the experiment.
The overall size of this dataset is 25GB.
The format of the data is \{ source url \} \{ some configuration \} \{ dest url \}
For the pagerank algorithm, only the source url and the dest url is needed.Therefore the dataset was was preprocessed to filter out the only the necessary information.

If the entire dataset is assumed to be represented by a huge graph, then here each DBPedia instance maps to a wikipedia article or a page which inturn could be mapped to a vertex of the graph. And the links between the instances could be mapped to  
edges of the graph.

\begin{figure}
\centering
  \includegraphics[width=80mm]{images/numVertices.png}
  \caption{The vertices count over various dataset size}
  \label{fig:numvertices}
\end{figure}

\begin{figure} 
  \centering
  \includegraphics[width=80mm]{images/numLinks.png}
  \caption{The average edge count per vertex for various dataset size}
  \label{fig:numlinks}
\end{figure}

The figure \ref{fig:numvertices} shows the variation in the no of vertices count for different dataset size which from ~40,000 nodes and the figure \ref{fig:numlinks} shows  the average number of links or edges per vertex for a smaller dataset is much higher as compared to the larger dataset. The smaller dataset are the first few DBPedia instances of the larger dataset. The first few pages probably have more links compared the pages in the end of the dataset and hence the average is linearly decreasing.

\emph{Algorithm:}
The basic idea of the page rank algorithm is to assign scores to the pages based on the following two aspects of the page.
	\begin{enumerate}
		\item {\centering pages with higher incoming links get high rank.}
		 \item {\centering pages with links from a high rank page also gets high rank.}
	\end{enumerate}

The pseudocode of the algorithm is as follows:
\begin{enumerate}
	\item {\centering Assign every page a rank of 1.}
	\item {\centering On each iteration have page p contribute to rankp/ |neighbours|p to its neighbors.}
	\item {\centering Set each page ranks to \(0.15 + 0.85  *  contributions\)}.
	\end{enumerate}

The scala implementation of this algorithm is as shown below:
\begin{lstlisting}[language=ML]
vallinks =// RDD of (url, neighbors) pairs
varranks =// RDD of (url, rank) pairs
for(i<- 1 to ITERATIONS) 
{
  valcontribs=links.join(ranks).flatMap{
  case (url, (links, rank)) => links.
  map( dest=> (dest, rank/links.size))
  }
  ranks =contribs.reduceByKey(_ + _).
  mapValues(0.15 + 0.85 * _)
}
ranks.saveAsTextFile(...) 
\end{lstlisting}

\emph{Bagel}:
The Bagel programming model was developed by spark.This is the implementation of the graph processing framework developed by google called the \emph{Pregel} system[??]. Bagel is still in the development stage where in more features are being added to the basic graph computation framework.

Any graph algorithm runs for several iterations and each iteration is called a Superstep.The graph is represented as a distributed dataset of $(K,V)$ pairs.  Here, the keys are vertex IDs, the values are vertex and associated states of the vertices.Every vertex in the graph will be associated with a user specified function.In each superstep, the algorithm runs the function associated with the vertex using the messages recieved from the previous superstep. The output is a new vertex state and a list of all outgoing messages which is then sent to all the other dependent vertices which is then used as input for the next iteration.

For the pagerank algorithm, as previously explained the vertices are pages and the edges are links.The list of messages generated at every iteration is the contribution of the pagerank for all its outgoing links.

Bagel provides APIs for vertices, edges and messages which could be extended to customize for any algorithm.This enables the user to write any compilcated code within few lines of code.

Further, bagel also provides combiner utility where in a reduction operation is applied on the outgoing message which reduces the communication overhead.The aggregator utility helps in reducing the message across all nodes for all vertices of the graph after every superstep.

\emph{Hash Partitioner}:The algorithm makes use of a hash partitioner which is a Controlled data partitioning utility that reduces communication cost.Also, it keeps the information of how data was partitioned across jobs for further reduction of communication cost.

\emph{Kyro serialization}: For the bagel program a special form of serialization technique was used, this  serializes the object sent over the network that will be cached in serialized form. The kryo serialization is much faster than the default java serialization used in the spark system as it the communication cost.
